---
title: "Assignment 7 – Working with XML and JSON in R"
author: "Cliff Lee"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(tibble)
```

## HTML, JSON and XML

This assignment involves reading three files in different formats into an R markdown file. The files were created manually for this exercise, although on the web, they probably exist in a raw form.

The chosen books are below:

```{r}

books <- tibble(
  Title = c("Take the Cannoli: Stories from the New World", "Gig: Americans Talk about Their Jobs", "Kitchen Confidential: Adventures in the Culinary Underbelly"),
  Author= c("Sarah Vowell","John Bowe, Marisa Bowe & Sabin Streeter", "Anthony Bourdain"),
  Format= c("HTML", "XML", "JSON")
)

kable(books) %>%
  kable_styling(latex_options = "scale_down")

```


Each book will be read from a source file into a local variable.

## HTML - Take the Cannoli: Stories from the New World

## XML - Gig: Americans Talk about Their Jobs

## JSON - Kitchen Confidential: Adventures in the Culinary Underbelly

