---
title: "Week 10 - Sentiment Analysis"
author: "Cliff Lee"
date: "10/30/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Brief

This assignment focuses on sentiment analysis from a couple of different corpora. First, an example analysis is reproduced from "Text Mining with R" (Robinson, J. S. and D.) using the novels of Jane Austen. Later, I apply similar sentiment analysis on Jack London's *White Fang* using a different sentiment lexicon.

# Sentiment Analysis with Jane Austen Novels

As an example of R's capabilities, the below code block from Chapter 2 (Robinson, J. S. and D.) shows how to perform sentiment analysis with Jane Austen's novels. The below code shows how to extract positive and negative lexicons from three different dictionaries: afinn; bing; and nrc.


```{r}

library(tidytext)
library("wordcloud")

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")


library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)


nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)


library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)


library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```


```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)



bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")


get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)


get_sentiments("bing") %>% 
  count(sentiment)

bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts


bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
	 y = NULL)


custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                                      stop_words)

custom_stop_words




tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))


library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
			  max.words = 100)

p_and_p_sentences <- tibble(text = prideprejudice) %>% 
unnest_tokens(sentence, text, token = "sentences")

p_and_p_sentences$sentence[2]



austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())


bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()



```


# More Analysis with *White Fang*

As another example, we'll perform similar sentiment analsyis on Jack London's *White Fang.*  For a different sentiment, we'll search for fear since this story takes place in a dramatic natural setting.


```{r}

library("gutenbergr")
library("readr")
library("tidyverse")
library("reshape2")


white_fang <- gutenberg_download(c(910))

text_book <- tibble(line=1:nrow(white_fang), white_fang$text)

colnames(text_book) <- c('line', 'text')
tidy_book <- text_book %>% unnest_tokens(word, text) %>% arrange(desc(line))

sentiment_positive <- get_sentiments("loughran") %>% 
  filter(sentiment == "positive")

sentiment_fear <- get_sentiments("nrc") %>% 
  filter(sentiment == "fear")

# get by negative sentiment
tidy_book %>%
  inner_join(sentiment_fear) %>%
  count(word, sort=TRUE)
```


Wordcloud for all words in *White Fang.*


```{r}
# make a word cloud
tidy_book %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

```

```{r}
# make a word cloud
tidy_book %>%
    inner_join(sentiment_fear) %>%

  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

```


The below section repeats the analysis with the bing lexicon of categorizing words as negative or positive.

```{r}
# get all sentiments
bing_word_counts <- tidy_book %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort=TRUE)

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ggplot(aes(reorder(word, n), n, fill = sentiment)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment", x = NULL) +
  coord_flip()


```


# References

Robinson, J. S. and D. (n.d.). Sentiment analysis with Tidy Data *Text mining with R*. Retrieved October 30, 2021, from https://www.tidytextmining.com/sentiment.html. 

London, Jack. *White Fang*, Retrieved October 30, 2021 from https://gutenberg.org/ebooks/910.

