---
title: "TidyVerse Vignette"
author: "Cliff Lee"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{clifflee-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2021TIDYVERSE


Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.


## Topic 

To illustrate how to use TidyVerse packages, datasets about college tuition costs will be consumed, tidied and analyzed. Specifically, tuition costs in 2018 and how they compare to 2010.

Reference: https://data.ed.gov/dataset/college-scorecard-all-data-files-through-6-2020/resources 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Setup

To get started, we'll simply load the tidyverse package which includes many other sub-packages.

```{r setup, message=FALSE}
library (tidyverse)
```


## Readr Functions

The Readr package provides functions to read regular, table formatted data from local or web files. The **read_csv** function below reads a CSV file into a dataframe.


```{r}

csv_2010 <- "https://raw.githubusercontent.com/cliftonleesps/607_acq_mgt/main/tidyverse/MERGED2010_11_PP.small.csv"
csv_2018 <- "https://raw.githubusercontent.com/cliftonleesps/607_acq_mgt/main/tidyverse/MERGED2018_19_PP.small.csv"

na_fields <- c("", "NA", "NULL", "PrivacySuppressed")


# Use the readr::read_csv function
college_2010 <- read_csv(csv_2010, na = na_fields, show_col_types = FALSE)
college_2018 <- read_csv(csv_2018, na = na_fields, show_col_types = FALSE)
```


## Dplyr Functions

Dplyr makes it easy to manipulate and join dataframes on various conditions.  The functions used in this block: **inner_join; mutate; filter; arrange; select**.

```{r}
?dplyr

# Use the dplyr::inner_join function to join by Office of Postsecondary Education ID as primary key.
# Created by the Department of Education.
all_colleges <- inner_join(college_2010, 
                           college_2018,
                           by = "OPEID"
                           )


# create a new column diff_debt
# Filter out any rows where the diff_debt == NA
all_colleges <- all_colleges %>%
  mutate( diff_debt = GRAD_DEBT_MDN.y - GRAD_DEBT_MDN.x) %>%
  filter(!is.na(diff_debt) & !is.na(COSTT4_A.y)) %>%
  arrange(desc(COSTT4_A.y), desc(diff_debt))


# Select only a few columns (dplyr::select)
display_table <- all_colleges %>% select("INSTNM.y","COSTT4_A.y","GRAD_DEBT_MDN.y","diff_debt")

display_table <- display_table[1:500,]

# rename columns
colnames(display_table) <- c("Institution Name","Annual Tuition Cost (2018)","Median Debt at Graduation","Difference in Median Debt from 2010")

```

## GGPlot 

For plotting and presentation graphics, ggplot2 offers advantages over the basic R ggplot package. The main advantage is being able to add aspects of the graph one by one without having to call functions with many parameters.  The function used:  **ggplot; geom_polint; aes; theme_bw; labs**.

```{r}


# Use the ggplot function to display our results
ggplot(data = display_table) +
  geom_point(
    mapping = aes(x = `Annual Tuition Cost (2018)`, y = `Difference in Median Debt from 2010`)
  ) +
  theme_bw() +
  labs(
      title = "Annual Tuition (2018) Compared to Change in Median Debt",
      x = "Annual Tuition 2018 ($)", 
      y = "Change in Median Debt ($)")


```
